---
title: "AI Web Signals"
subtitle: "How /llms.txt, robots.txt, and Content Signals overlap"
description: "A scan of the Tranco Top 100,000 shows that AI discovery, crawler access, and content preferences are separate web signals."
format:
  html:
    fig-width: 10
    fig-height: 6
---

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(scales)
})

source("analysis/data.R")
source("analysis/summaries.R")
source("analysis/plots.R")

dir.create("results/tables", recursive = TRUE, showWarnings = FALSE)
dir.create("results/figures", recursive = TRUE, showWarnings = FALSE)
write_csv(
  key_metrics,
  "results/tables/key-metrics.csv",
  na = ""
)
write_csv(
  restriction_source_summary,
  "results/tables/restriction-rule-sources.csv",
  na = ""
)
write_csv(
  content_summary,
  "results/tables/content-signals.csv",
  na = ""
)
write_csv(
  overlap_summary,
  "results/tables/llms-restriction-overlap.csv",
  na = ""
)
write_csv(
  category_summary,
  "results/tables/category-summary.csv",
  na = ""
)

ggsave(
  "results/figures/01-llms-presence.png",
  llms_presence_plot,
  width = 10,
  height = 4,
  dpi = 320,
  bg = "white"
)
ggsave(
  "results/figures/02-category-llms-presence.png",
  category_llms_plot,
  width = 10,
  height = 6.4,
  dpi = 320,
  bg = "white"
)
ggsave(
  "results/figures/03-llms-restriction-overlap.png",
  overlap_plot,
  width = 10,
  height = 4.6,
  dpi = 320,
  bg = "white"
)
ggsave(
  "results/figures/04-category-agent-restrictions.png",
  category_restriction_plot,
  width = 10,
  height = 6.4,
  dpi = 320,
  bg = "white"
)
ggsave(
  "results/figures/05-restriction-rule-source.png",
  restriction_source_plot,
  width = 10,
  height = 3.8,
  dpi = 320,
  bg = "white"
)
ggsave(
  "results/figures/06-content-signals.png",
  content_signals_plot,
  width = 10,
  height = 5,
  dpi = 320,
  bg = "white"
)

observed_llms_metric <- key_metrics |>
  filter(metric == "Observed /llms.txt among resolved observations")

explicit_policy_metric <- key_metrics |>
  filter(metric == "Explicit tracked-agent robots policy")

complete_scan_metric <- key_metrics |>
  filter(metric == "Complete endpoint scans")

resolved_llms_metric <- key_metrics |>
  filter(metric == "Resolved /llms.txt observations")

resolved_policy_metric <- key_metrics |>
  filter(metric == "Resolved tracked-agent policy observations")

llms_overlap <- overlap_summary |>
  filter(
    llms_txt == "Present",
    restriction == "Any tracked-agent restriction"
  )

category_llms_high <- category_plot_data |>
  slice_max(observed_llms_txt_share, n = 1, with_ties = FALSE)

category_llms_low <- category_plot_data |>
  slice_min(observed_llms_txt_share, n = 1, with_ties = FALSE)

category_restriction_high <- category_plot_data |>
  slice_max(any_agent_restriction_share, n = 1, with_ties = FALSE)

category_restriction_low <- category_plot_data |>
  slice_min(any_agent_restriction_share, n = 1, with_ties = FALSE)
```

AI discovery, crawler access, and content preferences are separate signals.
They can appear together, conflict, or remain unspecified. None establishes
permission, compliance, or legal status.

## `/llms.txt` remains uncommon

```{r}
#| label: llms-presence
#| echo: false
#| fig-height: 4
#| fig-alt: "A stacked bar comparing resolved domains with and without an observed llms.txt file."

llms_presence_plot
```

A plausible file appeared on
`r percent(observed_llms_metric$proportion, accuracy = 0.1)` of resolved
observations (`r comma(observed_llms_metric$numerator)` of
`r comma(observed_llms_metric$denominator)`).

### Adoption varies by category

Among categories with enough resolved observations to compare,
`r category_llms_high$category` had the highest observed rate at
`r percent(category_llms_high$observed_llms_txt_share, accuracy = 0.1)`
(`r comma(category_llms_high$observed_llms_txt)` of
`r comma(category_llms_high$resolved_llms_txt)`).
`r category_llms_low$category` was lowest at
`r percent(category_llms_low$observed_llms_txt_share, accuracy = 0.1)`.

```{r}
#| label: category-llms-presence
#| echo: false
#| fig-height: 6.4
#| fig-alt: "A ranked dot plot showing llms.txt presence rates for ten domain categories with at least 90 resolved observations. Software and technology is highest and government and public sector is lowest."

category_llms_plot
```

These are exploratory ChatGPT-assigned categories, not a validated taxonomy.
The comparison omits “Other / Unknown” and categories with fewer than
`r comma(category_min_resolved)` resolved observations for either signal.

## `/llms.txt` is not permission

Among domains where both endpoints resolved,
`r percent(llms_overlap$proportion, accuracy = 0.1)` with a plausible
`/llms.txt` also had a partial or full `robots.txt` restriction applicable to
at least one tracked agent (`r comma(llms_overlap$count)` of
`r comma(llms_overlap$denominator)`).

```{r}
#| label: discovery-restriction-overlap
#| echo: false
#| fig-height: 4.6
#| fig-alt: "Two bars show that 83.5 percent of domains with llms.txt and 66.9 percent without llms.txt restricted at least one tracked agent."

overlap_plot
```

This is a resolved-subset comparison, not evidence of association across all
`r comma(nrow(domains))` domains.

### Restrictions also vary by category

`r category_restriction_high$category` had the highest restriction rate among
the plotted categories at
`r percent(category_restriction_high$any_agent_restriction_share, accuracy = 0.1)`;
`r category_restriction_low$category` was lowest at
`r percent(category_restriction_low$any_agent_restriction_share, accuracy = 0.1)`.
The range is descriptive and may partly reflect which endpoints resolved.

```{r}
#| label: category-agent-restrictions
#| echo: false
#| fig-height: 6.4
#| fig-alt: "A ranked dot plot of applicable tracked-agent restriction rates for ten domain categories. News and publishing and e-commerce and retail are highest; cloud infrastructure is lowest."

category_restriction_plot
```

## Most restrictions are inherited

Only `r percent(explicit_policy_metric$proportion, accuracy = 0.1)` of resolved
domains explicitly addressed at least one tracked AI token. Most applicable
restrictions instead came from general `User-agent: *` rules.

```{r}
#| label: restriction-rule-source
#| echo: false
#| fig-height: 3.8
#| fig-alt: "A stacked bar shows that 90.5 percent of restrictive domain-agent observations came from wildcard rules and 9.5 percent from explicit agent rules."

restriction_source_plot
```

“Applicable” does not mean explicitly targeted. “Restricted” includes partial
path restrictions as well as full blocks.

## Content Signals remain rare

Content Signals express preferences, not access rules, and were rarely
specified when `robots.txt` could be interpreted.

```{r}
#| label: content-signals
#| echo: false
#| fig-height: 5
#| fig-alt: "A dot plot compares explicit yes and no Content Signal declarations. AI training refusals reach 5.7 percent while all other values remain below 1 percent."

content_signals_plot
```

## Coverage

```{r}
#| label: coverage-table
#| echo: false

tibble::tibble(
  Observation = c(
    "Both endpoints resolved",
    "/llms.txt resolved",
    "Tracked-agent policy resolved"
  ),
  Resolved = c(
    complete_scan_metric$numerator,
    resolved_llms_metric$numerator,
    resolved_policy_metric$numerator
  ),
  Unresolved = nrow(domains) - c(
    complete_scan_metric$numerator,
    resolved_llms_metric$numerator,
    resolved_policy_metric$numerator
  ),
  Coverage = c(
    complete_scan_metric$proportion,
    resolved_llms_metric$proportion,
    resolved_policy_metric$proportion
  )
) |>
  gt() |>
  tab_header(
    title = md("**Resolution is the study's main constraint**"),
    subtitle = paste0(
      "Resolved observations, not all ",
      comma(nrow(domains)),
      " selected domains, define outcome denominators."
    )
  ) |>
  fmt_integer(columns = c("Resolved", "Unresolved")) |>
  fmt_percent(columns = "Coverage", decimals = 1) |>
  cols_label(Coverage = paste0("Coverage of ", comma(nrow(domains)))) |>
  cols_align(align = "left", columns = "Observation") |>
  cols_align(align = "right", columns = c("Resolved", "Unresolved", "Coverage")) |>
  tab_style(
    style = cell_text(weight = "bold", color = "#2F6FA3"),
    locations = cells_body(columns = "Coverage")
  ) |>
  tab_source_note(
    "Endpoint failures may be systematic; resolved results are not estimates for unresolved domains."
  ) |>
  tab_options(
    table.font.names = "sans",
    table.font.size = px(15),
    data_row.padding = px(7),
    heading.align = "left",
    table.border.top.color = "#2F6FA3",
    table.border.top.width = px(3),
    column_labels.font.weight = "bold"
  )
```

Unresolved observations remain unknown. They are never recoded as absent,
allowed, or unrestricted.

## Method

The collector requested only `/llms.txt` and `/robots.txt` for the top
`r comma(nrow(domains))` pay-level domains in
[Tranco list 645ZX](https://tranco-list.eu/list/645ZX). The scan ran on
16 August 2026.

Rules cover training/control, AI search, and user-triggered fetch agents.
Results are descriptive counts and proportions with no weighting, imputation,
or inferential tests. Category figures use the exploratory labels in
`data/processed/categorized-domains.csv`. Rendering writes CSVs and PNGs to
`results/`.

## Limits

- Published signals do not prove enforcement, compliance, or permission.
- Wildcard and explicit rules are distinguished; partial rules are not full
  blocks.
- `/llms.txt` presence does not measure semantic quality.
- Category labels are model-assigned, exploratory, and unevenly distributed.
- Endpoint resolution was low and may be non-random.
- This is a time-bound scan of domains, not a random sample of websites.

## Citation

Victor Le Pochat, Tom Van Goethem, Samaneh Tajalizadehkhoob, Maciej
Korczyński, and Wouter Joosen. 2019. “Tranco: A Research-Oriented Top Sites
Ranking Hardened Against Manipulation,” *Proceedings of the 26th Annual Network
and Distributed System Security Symposium (NDSS 2019).*
<https://doi.org/10.14722/ndss.2019.23386>
