---
title: "AI Web Signals"
subtitle: "Observable AI Policy Signals Across the Tranco Top 100,000"
format:
  html:
    fig-width: 10
    fig-height: 6
---

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(purrr)
  library(readr)
  library(scales)
})

source("analysis/data.R")
source("analysis/summaries.R")
source("analysis/plots.R")

domains <- load_domains()
agent_policies <- load_agent_policies(domains = domains)
tranco_metadata <- load_tranco_metadata()

scan_quality <- summarise_scan_quality(domains)
endpoint_statuses <- summarise_endpoint_statuses(domains)
key_metrics <- summarise_key_metrics(domains)
rank_summary <- summarise_rank_bands(domains)
group_summary <- summarise_group_restrictions(domains)
agent_summary <- summarise_agent_policies(agent_policies)
content_summary <- summarise_content_signals(domains)
overlap_summary <- summarise_llms_restriction_overlap(domains)

llms_rank_plot <- plot_llms_by_rank(rank_summary)
group_restriction_plot <- plot_group_restrictions(group_summary)
agent_restriction_plot <- plot_agent_restrictions(agent_summary)
content_signal_plot <- plot_content_signals(content_summary)
overlap_plot <- plot_llms_restriction_overlap(overlap_summary)

table_outputs <- list(
  "key-metrics.csv" = key_metrics,
  "scan-quality.csv" = scan_quality,
  "endpoint-statuses.csv" = endpoint_statuses,
  "rank-bands.csv" = rank_summary,
  "group-restrictions.csv" = group_summary,
  "agent-policies.csv" = agent_summary,
  "content-signals.csv" = content_summary,
  "llms-restriction-overlap.csv" = overlap_summary
)

dir.create("results/tables", recursive = TRUE, showWarnings = FALSE)
purrr::iwalk(
  table_outputs,
  ~ readr::write_csv(.x, file.path("results", "tables", .y), na = "")
)

save_publication_plot(llms_rank_plot, "01-llms-presence-by-rank.png")
save_publication_plot(group_restriction_plot, "02-restrictions-by-purpose.png")
save_publication_plot(agent_restriction_plot, "03-restrictions-by-agent.png", height = 9)
save_publication_plot(content_signal_plot, "04-content-signals.png")
save_publication_plot(overlap_plot, "05-llms-restriction-overlap.png")

get_metric <- function(name) {
  metric <- key_metrics |>
    filter(.data$metric == name) |>
    select(.data$numerator, .data$denominator, .data$proportion)

  if (nrow(metric) != 1) {
    stop("Expected exactly one key metric named: ", name, call. = FALSE)
  }

  metric
}

most_restricted_agent <- agent_summary |>
  filter(!is.na(.data$restriction_proportion)) |>
  slice_max(.data$restriction_proportion, n = 1, with_ties = TRUE)

complete_scan <- get_metric("Complete endpoint scans")
resolved_llms <- get_metric("Resolved /llms.txt observations")
observed_llms <- get_metric("Observed /llms.txt among resolved observations")
resolved_policy <- get_metric("Resolved tracked-agent policy observations")
training_restriction <- get_metric("Any declared training/control restriction")
search_restriction <- get_metric("Any declared AI search restriction")
user_fetch_restriction <- get_metric(
  "Any declared user-triggered fetch restriction"
)

complete_scan_pct <- percent(complete_scan$proportion, accuracy = 0.1)
resolved_llms_pct <- percent(resolved_llms$proportion, accuracy = 0.1)
llms_txt_pct <- percent(observed_llms$proportion, accuracy = 0.1)
llms_txt_fraction <- paste0(
  comma(observed_llms$numerator),
  " of ",
  comma(observed_llms$denominator)
)
resolved_policy_pct <- percent(resolved_policy$proportion, accuracy = 0.1)
training_restriction_pct <- percent(
  training_restriction$proportion,
  accuracy = 0.1
)
search_restriction_pct <- percent(
  search_restriction$proportion,
  accuracy = 0.1
)
user_fetch_restriction_pct <- percent(
  user_fetch_restriction$proportion,
  accuracy = 0.1
)
most_restricted_agent_names <- toString(as.character(most_restricted_agent$agent))
most_restricted_agent_phrase <- if (nrow(most_restricted_agent) == 1) {
  paste(most_restricted_agent_names, "had")
} else {
  paste(most_restricted_agent_names, "shared")
}
most_restricted_agent_pct <- percent(
  most_restricted_agent$restriction_proportion[[1]],
  accuracy = 0.1
)
most_restricted_agent_n <- comma(most_restricted_agent$resolved_results[[1]])

list_id <- if (is.null(tranco_metadata$list_id)) "26J39" else tranco_metadata$list_id
list_url <- if (is.null(tranco_metadata$list_url)) {
  "https://tranco-list.eu/list/26J39/1000000"
} else {
  tranco_metadata$list_url
}
generated_on <- if (is.null(tranco_metadata$generated_on)) {
  "16 August 2026"
} else {
  as.character(tranco_metadata$generated_on)
}
```

This descriptive report measures public signals across popular domains: whether
a plausible `/llms.txt` file was observed, what `robots.txt` rules apply to
tracked AI-related agents and control tokens, and whether a site-wide
`Content-Signal` declaration expresses a preference for search, AI input, or AI
training. These observations describe what a domain published when it was
scanned; they do not demonstrate crawler compliance, provider behavior, or
legal permission.

## Coverage and selected findings

The study selected `r comma(nrow(domains))` ranked domains. Both endpoint
observations were complete for `r complete_scan_pct` of them. The `/llms.txt`
result was resolved for `r resolved_llms_pct` of selected domains; within that
resolved subset, a plausible file was observed for `r llms_txt_pct`
(`r llms_txt_fraction`).

A tracked-agent policy result was resolved for `r resolved_policy_pct` of
selected domains. Within that subset, `r training_restriction_pct` had at least
one declared restriction for a training/control token. The corresponding
descriptive proportions were `r search_restriction_pct` for AI search agents and
`r user_fetch_restriction_pct` for user-triggered fetch agents. These are
resolved-subset descriptions—not estimates for all 100,000 domains—and combine
partial restrictions with full blocks.

```{r}
#| label: key-metrics

key_metrics |>
  select(
    Metric = .data$metric,
    Numerator = .data$numerator,
    Denominator = .data$denominator,
    Proportion = .data$proportion
  ) |>
  gt() |>
  fmt_integer(columns = c("Numerator", "Denominator")) |>
  fmt_percent(columns = "Proportion", decimals = 1) |>
  sub_missing(
    columns = c("Denominator", "Proportion"),
    missing_text = "—"
  ) |>
  tab_header(title = "Coverage and selected descriptive measurements") |>
  tab_source_note(
    paste(
      "Each proportion uses the denominator shown; unresolved observations",
      "are not recoded as negative results."
    )
  ) |>
  tab_options(table.font.names = "sans")
```

## Collection quality

Unresolved values are analytically distinct from negative observations. Missing
logical values remain missing, the categorical `unknown` state remains visible,
and no imputation or weighting is applied. Because endpoint failures may be
systematic, results for resolved observations should not be generalized to
unresolved domains.

```{r}
#| label: scan-quality

scan_quality |>
  transmute(
    `Scan status` = tools::toTitleCase(as.character(.data$scan_status)),
    Count = .data$count,
    Proportion = .data$proportion
  ) |>
  gt() |>
  fmt_integer(columns = "Count") |>
  fmt_percent(columns = "Proportion", decimals = 1) |>
  tab_options(table.font.names = "sans")
```

Endpoint classifications are retained separately because a partial scan can
still produce a usable result for one endpoint.

```{r}
#| label: endpoint-statuses

endpoint_statuses |>
  arrange(.data$endpoint, desc(.data$count)) |>
  mutate(
    endpoint = if_else(.data$endpoint == "llms.txt", "/llms.txt", "/robots.txt"),
    status = gsub("_", " ", .data$status)
  ) |>
  select(
    Endpoint = .data$endpoint,
    Status = .data$status,
    Count = .data$count,
    Proportion = .data$proportion
  ) |>
  gt(groupname_col = "Endpoint") |>
  fmt_integer(columns = "Count") |>
  fmt_percent(columns = "Proportion", decimals = 1) |>
  tab_options(table.font.names = "sans")
```

## Observed `/llms.txt` presence

Observed presence means the collector found a plausible public `/llms.txt`
file. It does not assess semantic quality, usefulness, or whether any AI
provider uses the file.

```{r}
#| label: llms-presence-by-rank

llms_rank_plot
```

## Declared `robots.txt` policies for tracked agents

The analysis separates training/control tokens, AI search/indexing agents, and
user-triggered fetch agents. `None`, `Some`, and `All` summarize how many
tracked agents or tokens in a purpose group have partial or full restrictions.
`Unknown` identifies unresolved policy observations; it is not counted as
unrestricted.

```{r}
#| label: restrictions-by-purpose

group_restriction_plot
```

At agent grain, `r most_restricted_agent_phrase` the largest observed
restriction proportion: `r most_restricted_agent_pct` among
`r most_restricted_agent_n` domains with a resolved policy for that token. This
describes published rules in the resolved subset, not whether the named agent
obeyed them.

```{r}
#| label: restrictions-by-agent
#| fig-height: 9

agent_restriction_plot
```

## Content Signals

`Content-Signal` declarations are separate from crawler access rules. `Yes`
and `No` are explicit preferences, `Unspecified` means no applicable site-wide
declaration was found in an otherwise interpretable robots observation,
`Invalid` preserves malformed or conflicting declarations, and `Unknown`
marks observations that could not be evaluated reliably. The figure uses all
selected domains as its denominator so that unresolved coverage remains visible.

```{r}
#| label: content-signals

content_signal_plot
```

## Observed `/llms.txt` and restriction signals

Publishing `/llms.txt` and declaring a restriction for a tracked agent are
separate choices. The following descriptive cross-tabulation includes only
domains for which both signals were resolved; proportions are calculated within
each `/llms.txt` row. It is not an estimate of association across all selected
domains.

```{r}
#| label: discovery-restriction-overlap

overlap_plot
```

## Methodology

We scanned the top 100,000 pay-level domains from the standard Tranco list for
publicly observable AI-related signals in `robots.txt` and `llms.txt`. The
current inventory uses Tranco list `r list_id`, generated on `r generated_on`.^[
Available at [`r list_url`](`r list_url`).] When
`data/input/tranco-metadata.json` is present, the report uses that file as the
source of truth for list provenance. The list was fetched and the scan was run
on 16 August 2026.

For each domain, the collector requested only `/llms.txt` and `/robots.txt`.
The domain-level dataset contains endpoint classifications, three site-wide
Content Signal results, explicit-policy and any-restriction flags, and grouped
restriction states. The agent-level dataset contains one row for each of 15
tracked agents per domain. Partial and full restrictions both count as
restrictive in proportion calculations, while full blocks remain available
separately in the exported agent table.

Tranco rank is an aggregate rank over pay-level domains. It should not be read
as traffic volume, exact visitors, page views, or HTTP reachability. Analysis
denominators distinguish all selected domains, resolved endpoint observations,
and determined AI-policy results. Failed network observations are not negative
AI-policy observations.

This is a descriptive census attempt for the selected Tranco list, not a random
sample of the web. The analysis reports counts and proportions without
inferential tests, confidence intervals, weighting, or missing-value imputation.
Rank-band comparisons are conditional on successful endpoint resolution within
each band.

The render writes reproducible summary CSVs to `results/tables/` and PNG figures
to `results/figures/`. The exact field definitions, tracked-agent list, and
collection limitations are documented in the repository README.

## Limitations

- `robots.txt` and Content Signals express declared policy or preference; they
  do not prove enforcement, compliance, or legal permission.
- User-agent strings can be spoofed, and user-triggered agents may behave
  differently from autonomous crawlers.
- `/llms.txt` is classified for plausible presence, not semantic quality.
- Partial restrictions do not imply that every path is blocked.
- Endpoint resolution is low and may be non-random. Resolved-subset proportions
  should not be interpreted as prevalence estimates for all selected domains.
- The population includes infrastructure and machine-facing domains as well as
  consumer-facing sites, so this report consistently refers to domains rather
  than websites.
- Results are a time-bound snapshot of live, mutable resources.

## Citation

Victor Le Pochat, Tom Van Goethem, Samaneh Tajalizadehkhoob, Maciej
Korczyński, and Wouter Joosen. 2019. "Tranco: A Research-Oriented Top Sites
Ranking Hardened Against Manipulation," *Proceedings of the 26th Annual Network
and Distributed System Security Symposium (NDSS 2019).*
<https://doi.org/10.14722/ndss.2019.23386>
