---
title: "AI Web Signals"
subtitle: "How /llms.txt, robots.txt, and Content Signals overlap"
description: "A scan of the Tranco Top 100,000 shows that AI discovery, crawler access, and content preferences are separate web signals."
format:
  html:
    fig-width: 10
    fig-height: 6
---

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(scales)
})

source("analysis/data.R")
source("analysis/summaries.R")
source("analysis/plots.R")

dir.create("results/tables", recursive = TRUE, showWarnings = FALSE)
dir.create("results/figures", recursive = TRUE, showWarnings = FALSE)
write_csv(
  key_metrics,
  "results/tables/key-metrics.csv",
  na = ""
)
write_csv(
  restriction_source_summary,
  "results/tables/restriction-rule-sources.csv",
  na = ""
)
write_csv(
  content_summary,
  "results/tables/content-signals.csv",
  na = ""
)
write_csv(
  overlap_summary,
  "results/tables/llms-restriction-overlap.csv",
  na = ""
)
write_csv(
  category_summary,
  "results/tables/category-summary.csv",
  na = ""
)

ggsave(
  "results/figures/01-category-comparison.png",
  category_comparison_plot,
  width = 10,
  height = 6.4,
  dpi = 320,
  bg = "white"
)
ggsave(
  "results/figures/02-llms-restriction-overlap.png",
  overlap_plot,
  width = 10,
  height = 4.2,
  dpi = 320,
  bg = "white"
)
ggsave(
  "results/figures/03-content-signals.png",
  content_signals_plot,
  width = 10,
  height = 4.4,
  dpi = 320,
  bg = "white"
)

observed_llms_metric <- key_metrics |>
  filter(metric == "Observed /llms.txt among resolved observations")

explicit_policy_metric <- key_metrics |>
  filter(metric == "Explicit tracked-agent robots policy")

complete_scan_metric <- key_metrics |>
  filter(metric == "Complete endpoint scans")

resolved_llms_metric <- key_metrics |>
  filter(metric == "Resolved /llms.txt observations")

resolved_policy_metric <- key_metrics |>
  filter(metric == "Resolved tracked-agent policy observations")

wildcard_source_metric <- restriction_source_summary |>
  filter(source == "Wildcard rule")
```

## By category

A plausible `/llms.txt` appeared on
`r percent(observed_llms_metric$proportion, accuracy = 0.1)` of resolved checks
(`r comma(observed_llms_metric$numerator)` of
`r comma(observed_llms_metric$denominator)`).

```{r}
#| label: category-comparison
#| echo: false
#| fig-height: 6.4
#| fig-alt: "Two aligned dot plots compare llms.txt presence and tracked-agent restrictions across ten adequately resolved domain categories. Software leads llms.txt presence, while news and retail lead restrictions."

category_comparison_plot
```

## `/llms.txt` is not permission

```{r}
#| label: discovery-restriction-overlap
#| echo: false
#| fig-height: 4.2
#| fig-alt: "Two bars show that 83.5 percent of domains with llms.txt and 66.9 percent without llms.txt restricted at least one tracked agent."

overlap_plot
```

This resolved-subset comparison is descriptive, not evidence of association.
`r percent(wildcard_source_metric$proportion, accuracy = 0.1)` of restrictive
domain-agent observations came from `User-agent: *`; only
`r percent(explicit_policy_metric$proportion, accuracy = 0.1)` of resolved
domains explicitly addressed a tracked AI token.

## Content preferences

```{r}
#| label: content-signals
#| echo: false
#| fig-height: 4.4
#| fig-alt: "A dot plot compares explicit Content Signal declarations. Search opt-ins reach 6.2 percent and AI training opt-outs reach 5.7 percent; the other values remain below 1 percent."

content_signals_plot
```

## Coverage

```{r}
#| label: coverage-table
#| echo: false

tibble::tibble(
  Observation = c(
    "Both endpoints resolved",
    "/llms.txt resolved",
    "Tracked-agent policy resolved"
  ),
  Resolved = c(
    complete_scan_metric$numerator,
    resolved_llms_metric$numerator,
    resolved_policy_metric$numerator
  ),
  Coverage = c(
    complete_scan_metric$proportion,
    resolved_llms_metric$proportion,
    resolved_policy_metric$proportion
  )
) |>
  gt() |>
  tab_header(
    title = md("**Only about one in ten endpoints resolved**")
  ) |>
  fmt_integer(columns = "Resolved") |>
  fmt_percent(columns = "Coverage", decimals = 1) |>
  cols_label(Coverage = paste0("Of ", comma(nrow(domains)))) |>
  cols_align(align = "left", columns = "Observation") |>
  cols_align(align = "right", columns = c("Resolved", "Coverage")) |>
  tab_style(
    style = cell_text(weight = "bold", color = "#2F6FA3"),
    locations = cells_body(columns = "Coverage")
  ) |>
  tab_source_note(
    "Endpoint failures may be systematic; resolved results are not estimates for unresolved domains."
  ) |>
  tab_options(
    table.font.names = "sans",
    table.font.size = px(15),
    data_row.padding = px(7),
    heading.align = "left",
    table.border.top.color = "#2F6FA3",
    table.border.top.width = px(3),
    column_labels.font.weight = "bold"
  )
```

## Method

The collector requested `/llms.txt` and `/robots.txt` for the 100,000 pay-level
domains in [Tranco list 645ZX](https://tranco-list.eu/list/645ZX) on 16 August
2026. It tracks training/control, AI search, and user-triggered fetch agents.

Results are descriptive, with no weighting, imputation, or inference.
Unresolved observations remain unknown; missingness may be systematic.
Categories are exploratory. Partial restrictions are not full blocks, and
published signals do not establish permission, compliance, enforcement, or use.

## Citation

Victor Le Pochat, Tom Van Goethem, Samaneh Tajalizadehkhoob, Maciej
Korczyński, and Wouter Joosen. 2019. “Tranco: A Research-Oriented Top Sites
Ranking Hardened Against Manipulation,” *Proceedings of the 26th Annual Network
and Distributed System Security Symposium (NDSS 2019).*
<https://doi.org/10.14722/ndss.2019.23386>
